Differential Expression - HBCC - dsACC - Dx

### 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
test DE HBCC dsACC

---
title: "Differential Expression - HBCC - sgACC - Dx"
author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
date: "`r date()`"
output: html_document
---

@@

knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_sgACC_DE.Rmd",
                                 parentId ="syn22084216",
                                 folderName = "Differential Expression - Dx")
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22084216';
activityName = 'Differential Expression';
activityDescription = 'Differential Expression - Dx';

thisFileName <- 'HBCC_sgACC_DE.Rmd'

@@

# Download CQN counts
CQN_ID = "syn22084946"
ALL_USED_IDs = c(CQN_ID)
CQN.GENE_EXPRESSION = downloadFile(CQN_ID) %>% 
  tibble::column_to_rownames(var = "ensembl_gene_id") %>% 
  data.matrix()

# Download filtered counts
COUNTS_ID = "syn22084938"
ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
NEW.COUNTS = downloadFile(COUNTS_ID) %>% 
  tibble::column_to_rownames(var = "ensembl_gene_id")

# Get covariates
COVAR_ID = "syn22084935"
ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
COVARIATES = downloadFile(COVAR_ID) %>% 
  tibble::column_to_rownames(var = "SampleID")

@@

# Set covariates to adjust for from model selection
adjust.covars <- c("Dx", "RIN2", "IntronicRate")
random_effect <- c("Reported_Gender", "LibraryBatch")
```

@@

# Model categorical variables as random to report variance fractions
formula_for_variance_fractions = sapply(1:length(col_type$class), function(i){
  switch(col_type$class[i], 
    "factor" = paste0('(1|', col_type$variable[i], ')'),
    "numeric" = paste0(col_type$variable[i])
  )
  })

@@

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
  dplyr::mutate(Tissue = "sgACC")

@@

# Fit contrast
FIT.CONTR = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
                                     formula = formula,
                                     data = COVARIATES,
                                     L = L,
                                     suppressWarnings = TRUE)

@@

# Get gene lengths
counts = 'syn21867762'
ALL_USED_IDs = counts
counts = downloadFile(counts) %>% data.frame()
gene_lengths = counts[,c("Geneid", "Length")] %>% 
  dplyr::rename(gene_id = Geneid)
rm(counts)

@@

# Store residual gene expression for network analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_sgACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_sgACC_Dx_netResidualExpression.tsv', 
                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(nEXP_OBJ, annotations = all.annotations)

# Store differential expression results
write.table(DE, file = 'CMC_sgACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('CMC_sgACC_Dx_DiffExpression.tsv', 
                name = 'Differential Expression Results (Dx)', 
                parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(DEXP_OBJ, annotations = all.annotations)

# Store adjusted fit linear mixed model object as RDS - Dx
saveRDS(ADJUSTED.FIT, file = 'CMC_sgACC_Dx_dream-output.rds')
ADJ_OBJ = File('CMC_sgACC_Dx_dream-output.rds',
               name = 'Dream Output object - Dx', 
               parentId = CODE$properties$id)
ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
                   executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
```

>>>>

---
title: "Differential Expression - HBCC - dsACC - Dx"
author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
date: "`r date()`"
output: html_document
---

@@

knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_DE.Rmd",
                                 parentId ="syn22084347",
                                 folderName = "Differential Expression - Dx")
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22084347';
activityName = 'Differential Expression';
activityDescription = 'Differential Expression - Dx';

thisFileName <- 'HBCC_dsACC_DE.Rmd'

@@

# Download CQN counts
CQN_ID = "syn22086076"
ALL_USED_IDs = c(CQN_ID)
CQN.GENE_EXPRESSION = downloadFile(CQN_ID) %>% 
  tibble::column_to_rownames(var = "ensembl_gene_id") %>% 
  data.matrix()

# Download filtered counts
COUNTS_ID = "syn22086075"
ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
NEW.COUNTS = downloadFile(COUNTS_ID) %>% 
  tibble::column_to_rownames(var = "ensembl_gene_id")

# Get covariates
COVAR_ID = "syn22086074"
ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
COVARIATES = downloadFile(COVAR_ID) %>% 
  tibble::column_to_rownames(var = "SampleID")

@@

# Set covariates to adjust for from model selection
adjust.covars <- c("Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
random_effect <- c("Reported_Gender", "IsolationBatch")
```

@@

# Model categorical variables as random to report variance fractions
col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
  dplyr::summarise_all(class) %>% 
  tidyr::gather(variable, class)

formula_for_variance_fractions = sapply(1:length(col_type$class), function(i){
  switch(col_type$class[i], 
    "factor" = paste0('(1|', col_type$variable[i], ')'),
    "numeric" = paste0(col_type$variable[i])
  )
  })

@@

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
  dplyr::mutate(Tissue = "dsACC")

@@

# Fit contrast
FIT.CONTR = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,
                                     formula = formula,
                                     data = COVARIATES,
                                     L = L,
                                     suppressWarnings = TRUE)

@@

# Get gene lengths
counts = 'syn21867866'
ALL_USED_IDs = counts
counts = downloadFile(counts) %>% data.frame()
gene_lengths = counts[,c("Geneid", "Length")] %>% 
  dplyr::rename(gene_id = Geneid)
rm(counts)

@@

# Store residual gene expression for network analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_dsACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_dsACC_Dx_netResidualExpression.tsv', 
                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(nEXP_OBJ, annotations = all.annotations)

# Store differential expression results
write.table(DE, file = 'CMC_dsACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('CMC_dsACC_Dx_DiffExpression.tsv', 
                name = 'Differential Expression Results (Dx)', 
                parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(DEXP_OBJ, annotations = all.annotations)

# Store adjusted fit linear mixed model object as RDS - Dx
saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Dx_dream-output.rds')
ADJ_OBJ = File('CMC_dsACC_Dx_dream-output.rds',
               name = 'Dream Output object - Dx', 
               parentId = CODE$properties$id)
ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
                   executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
```
